\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{cite}

\title{Programming as a teaching tool in third-grade mathematics }
\author{Eric Bagley}
\date{June 2020}

\begin{document}

\maketitle

\section*{Introduction}

Transitions move us from where we are to what comes next.
They can be seamless or more difficult depending on the situation.
Elementary school is full of those transitions.
Some students maneuver through these transitions easily while other students find them much more difficult. 
In the latter situation teachers often need to be creative in order to find ways to help facilitate these transitions.

Learning the basics of mathematics plays a key role in developing logical and abstract thinking skills and problem-solving skills.
For some students, mathematical concepts are simple; they just make sense.
For others, these concepts don’t come easily at all.
These preconceptions often originate from a number of reasons but that is a conversation outside the focus in this paper.
The fact is, understanding the complexities of mathematics is difficult for many students to understand.

% Brief Introduction of the education standards
One of the most challenging parts of education is determining what to teach and how to teach it.
Some recent attempts at education reform include the No child left behind Act,  the Every Student Succeeds Act and the Common Core Standards.
Many education reforms, or at least parts of them, introduce positive and promising results while others perform less than optimal.
Regardless of what system is in place, there is a need for teaching tools that help students develop skills that will help them in all aspects of life.
Creativity, problem-solving, critical thinking, and communication are vital for someone to become well rounded, responsible individuals that contribute to their home-life, work, and communities.

% Brief Introduction of third graders
The focus of the proposed study is not to propose a system that outlines education reform nor to argue that there is a need for it.
Its focus is much more specific: third-grade mathematics.
Third-grade is a common time for students to begin learning about multiplication and division, fractions, structures of arrays and area, and describing and analyzing two-dimensional shapes, place value, patterns, and problem-solving, number operations, and measurement and data analysis.
The key principles behind learning these concepts are to make sense of problems and attempt to solve them, reason abstractly and quantitatively, construct viable arguments and critique the reasoning of others, model with mathematics, and to look for patterns.

% Where are they in their cognition development?
By age 8, children start valuing relationships more, think about the future, and become more independent from parents and family.
They can more accurately describe their feelings and understand more about their place in the world.
They are beginning to see and understand things from another person’s point of view.
Their ability to solve problems increases and their attention span gets longer.
In math at school, they should be pretty comfortable with addition and subtraction, and they are becoming more proficient with multiplication.

% Why is the current teaching method insufficient
Technology is increasingly becoming a more integral part of our daily lives.
One might argue that teaching multiplication and division are no longer necessary because technology can perform those operations for us.
Although the latter statement may be true students benefit from skills developed while learning those mathematical operations.
Teaching these topics, or at least the principles behind them, are ever more important because technology often does much of the thinking for us.
Integrating technology and computational thinking will be extremely valuable to students in their current and future everyday life.

% TODO: Source this
One of the major transitions from second-grade to third-grade mathematics is moving from array-based thinking, using addition, subtracted, repeated addition, and repeated subtraction, to a unit-based, more abstract, thinking, using multiplication and division.
%[SOURCE].
This later leads to fractions and ratios.
This transition is often difficult for students and a non-traditional approach may prove beneficial.

% TODO
Another major goal of mathematics at this age is to help the students understand the fundamental pieces of numerical operations and how they are related to one another.
Although quick answer algorithms exist, without a strong understanding of the functions behind them it is difficult for students to truly understand the concept.
If this is the case, students often just plug in the numbers and hope that they turn out.
This may provide correct solutions, but it doesn't lend itself well to solving problems that haven't already been solved.
This is detrimental to subsequent learning because every step builds on the previous.
If this happens, later on, teachers find that they need to fill in the gaps.
This could introduce remedial coursework that may not be needed by other students and may cause disinterest or lost opportunities for deeper learning.

% What are the benefits that could come from this study 
Understanding the aforementioned misconceptions and shortcomings offer us key insights that can help us develop different teaching tools that help students fully understand fundamental concepts.
We propose using programming as a method for teaching multiplication and division at a third-grade level as it could provide an appropriate transition from the group based thinking students have developed from second-grade to the unit based, more abstract thinking they need to develop for future problem-solving skills.
It has the potential to assist students in understanding fundamental concepts and fill learning gaps before they become detrimental to both the individual student’s learning and the learning of their classmates.
It offers an appropriately tailored transition using programming to help them solidify the basics and move to solve more advanced topics.

% What type of study we hope to carry out (quantitative and qualitative)
This study will have a control group and a test group involving two parts, quantitative analysis, and qualitative analysis.
The qualitative portion is built off a pre-test and post-test.
The qualitative analysis will consist of various pieces: recordings of thinking out loud, pre- and post-interviews, and observation of classroom instruction and participation.


\section*{Related Work}
%When did programming as a teaching method start?
%Who pioneered it and what was their main objective?
%How has it evolved through the years?
\cite{sherin}
%What research in this area has taken place in the last 10 years?
%What is the main focus of this research?
%What are the findings of this research?
%Are there still unanswered questions? What are the areas of Further study mentioned in the papers?

Using programming as a teaching method for topics outside of computer science is not novel.
There have been many studies related to pedagogy and programming across all educational levels.
Within each level, a wide range of topics are covered.
Computer Science principles are used, however, to facilitate the learning process.
These principles include computational thinking, problem-solving, planning, and creativity.
The aforementioned principles have been used to teach language, writing, mathematics, physics, and other science topics.

% \subsection*{Learning to Code}
The studies reviewed can be placed into three categories: ones that teach computer science principles, ones that teach programming languages, and those that use computer science principles and programming languages to teach subjects outside of computer science.

% TODO add source
The first area addressed in a number of studies have research objectives that focus on determining appropriate ways to teach a specific computer science principle.
Other studies focus on determining appropriate ways to teach programming syntax such as Phanon [EDWARDS].
Based on the target audience there is a variety of teaching software and a variety of different languages that aim to teach different principles.
For example, block-based visual programming targets younger or more novel students, mixed block-text based programming languages such as Droplet can be used for older or familiar students, and text based languages are more easily understood by secondary and post-secondary students.
Each of these have varying levels of scaffolding.
Block programs heavily utilize scaffolding by providing hints and guiding programmers develop code using close to natural language while text based languages can have little to no scaffolding.

Some programs aim to provide dynamic scaffolding that offers more intensive supports initially but gradually removes some or all scaffolding support for more advanced users.
    
\subsection*{Learning through Code}
The second area addressed in a number of studies is using software programs or programming languages as tools to teach educational subjects outside of computer science.
Different than virtual manipulatives or other computer games, programming has proved to be beneficial in helping students develop problem-solving, planning, meta-cognition and other skills \cite{papadakis, fuchs}.

\subsection{Coding blocks}\cite{}
\subsection{Programming Physics}\cite{sherin}

\subsection{Science Topics} \cite{guzdial}
\subsection{Story telling and Language} \cite{}
\subsection{Mathematics} \cite{}
\subsection{Virtual Manipulatives}\cite{lye}

It is widely known that programming can be used as a successful teaching tool.
There have been several studies focused on teaching mathematics, many of them use programming to do such.
The age group of said studies ranges from preschool to post-secondary.

\section*{Idea}
% Why is this study relevant to me? Why is it important to me?
Technology is an integral piece of our lives. Digital literacy is vital in everyday life.
I am a father; I want my children to have a great education.
I have friends who have struggled with math their whole life and I have friends that don't want anything to do with code or math. I think this is detrimental to them because they don't understand that piece of the world around them. 
% What Do I want to accomplish?
I want to help integrate computational thinking into everyday curriculum.
I want to determine for myself if programming can be a viable tool for teaching.
% What questions do I have?
\subsection*{Research Questions}
\begin{enumerate}
	\item Does programming math provide better transitions for third graders?
	
	I think that it could. As one paper said, programming isn't difficult it just requires clear thinking. 

	\item Is there a better way to teach math (using programming or the more traditional method)? 
	
    I believe that there needs to be deeper understanding. From my conversations with educators it seems that the goal is the same. There have been education reforms with hopes to improve it for all students from primary to post secondary. There is a strong desire to improve but there are some road blocks. 
    
		Those road blocks, I believe, are time limitations; fulfilling a number of requirements imposed by federal, state, and local governments; diverse class sizes, capabilities, and students; lack of funding;
		Many of these roadblocks can be remedied, while others require more thorough investigation, others may not be feasible to change. Lack of funding is difficult but it can be worked around in creative ways. Divers classrooms can be both beneficial and challenging, providing this type of learning activity may be great for some kids but more challenging for others. Time limitations are difficult to work around but again with creativity hopefully we can develop more effective teaching tools to decrease the amount of time required to teach students, decrease the amount of time it takes for students to learn the concepts, and decrease the amount of time teachers need to put toward reviewing and evaluating performance.
	
	\item Can programming be used for more than creating games and telling stories at this age?
	
	I believe it can. Using appropriate levels of scaffolding, programming can be a useful tool to help in specific areas of teaching. As Sherin suggests
	
	\begin{quotation}
			The point here is that programming forces a student to engage with these issues in a way that algebra-physics does not. … it is not too surprising that algebra-physics instruction leaves some undiscovered intuitive cobwebs in this territory, and that the act of programing tends to expose these cobwebs 	p48\cite{sherin}
	\end{quotation}
	\begin{quotation}
		...programming physics draws more strongly on causal intuitions	p50\cite{sherin}
	\end{quotation}
	\begin{quotation}
		...programs might be easier to understand or interpret than equations. 	3\cite{sherin}
	\end{quotation}
	\begin{quotation}
		...the act of programming requires the students to explode each instant of the motion into a series of actions that happen through what I will refer to as ‘pseudo-time’	34\cite{sherin}
	\end{quotation}

	\item How can we better incorporate digital literacy into our education?
    
    Using technology in the classroom helps students become digitally literate, but being able to tell a computer what to do increases said literacy. 
    
\end{enumerate}
\subsection{Questions}

% Are there any other questions that we might investigate?

\section*{Study}
% When will it happen? What semester? 
We hope to carry out the study during spring semester of 2021.
% What does the study Look like?
% How long will the study be? How many classroom hours will be required? Will it involve home study? 
The study will take place over the course of a few weeks
% How would this study look if it were done in a group face to face setting?
In an in person classroom setting there could be many ways to organize this study. 
One - Teacher lead:
The teacher can lead the students in the programming activity. This may be helpful first starting out to get the general idea of what the programming environment is like and students can ask questions.
Two - Student Lead, Teacher Oversee: 
The teacher can select students to lead the class to solve a certain problem or complete a step. After the problem or step is complete another student can control the computer. Students not controlling can work together to help the controller to write the code.(similar to ladybug programming in \cite{fessakis}).
% How would this study look if it were done individually online?
In an online format with a single user, the program would look similar to the in person version. However, in this case, as student would carry out the exercises by her/himself. 
% Which is preferred or better? Group activities vs individual? (in what ways are each stronger/weaker?)
Both individual and classroom exercises could provide different learning focuses. 

In a classroom setting children may give different facial or verbal queues that indicate the level of their understanding. The teacher could also help students at varying levels of achievement succeed by controlling the activities assigned and which solely activities are observed by each student. It may provide students the opportunity to learn from the experiences of their classmates. A disadvantage of classroom group activities is that each student may not reflect or communicate their actual understanding. This may lead the teacher to think they are either understanding more than they actually are, or that they are understanding less than they display. This may also diminish the motivation of some students as they may not feel comfortable in group settings. 

In an individual setting one advantage is that the activity can be tailored to the individual student. This may be more beneficial for lower performing students as it limits distractions, embarrassing moments in front of their classmates, and allows them to perform on their own timeline. Observing in this situation may be more difficult and the data could be more limited than in the in person setting. 

% What type of compensation will the participants receive?
A small amount of funds will be dedicated to the students, teachers, and researchers, participating in the study. 

% Will teachers need to put much more effort into their class prep?
Recognizing that teachers have limited time to cover materials, one goal of the study is to limit the amount of extra time needed for these lessons by making them easy to incorporate in their current curriculum. This will involve consistent and thorough communication with the teachers as to the learning objectives, general time dedicated to such objectives, and evaluation periods. 
The programming environment should not only have a simple interface that is familiar to students, but also provide appropriate levels of transparency and visibility into students' performance that help educators make appropriate evaluation decisions.

% What technological resources would we have to work with? What assumptions can we make? 
We are assuming that all of the students will have some access to a computer and the internet. This may be on a mobile device (preferably a tablet) or a computer with or without a touch screen.

% What type of data will we get from the study?
\section*{Data}
\begin{enumerate}
    \item Timing data: how long does it take each student to complete an activity.
    \item How many of the tutorials, hints, and other help and scaffolding methods were used by the student. 
    \item Mouse Movements or touch points (possibly, this could tell us what their program usage looks like.)
    \item How many blocks did they use?
    \item How many blocks did they drag on? 
    \item How many blocks did they throw away?
    \item What did their code look like? 
    \item What was the running time of their code?  
    \item Information about there devices can also useful as research shows that touch screen devices provide richer learning experiences.
    \item Basic demographic data (ethnicity, household income, do both parents work in or outside of the home)
    \item Previous education performance of students (this could just be the pre-test).
    \item Pre-evaluation.
    \item Post evaluation.
    \item Verbal feedback during learning activities
    \item Touch or mouse feedback
    \item Questionnaire (about the activity: What did you learn? Was it fun? Why? Do you understand [INSERT MATH CONCEPT] more than before? Explain what [INSERT MATH CONCEPT] means? etc)
\end{enumerate}

\section*{Participants}
% Who will participate in the study?
Students, Teachers, Researchers, Student's Parents (possibly)
% How will students participate in the study?
The students will participate in the programming activities. They may carry them out individually, as a class, or in small groups. They will provide feedback about the programming environment, learning activity, and their general experience.

% How will teachers participate in the study?
Teachers will introduce the course material and the learning activities. They may also introduce the programming environment. They help observe the students and their progress. They may provide feedback about the study, the programming environment, the learning activities, and their general experience. The teacher works with the researchers to develop the programming environment and learning activities for the students.

% How will I and other researchers participate in the study?
Researchers may provide feedback about the study. With advising from teachers, they prepare the programming environment and learning activities. They also analyze the data and report findings.

% In what ways will teachers be compensated?
% In what ways will students be compensated?
% In what ways will researchers be compensated?

\section*{Programming Environment}
% What Programming language did we choose to use?
% What were some programming languages we looked into but chose not to use?
% Why did we choose the Language we did over the others?
% What are the benefits from using this programming environment?
% What are the risks that come from using this programming environment?
% How much time will be required on my part to develop this programming environment?

\section*{Math Topics}
% What Math topics could be covered in the study?
% How might the programming activity look?
% How will we determine if math is being learned in each activity? What evidences will lead us to believe it is happening?

\section*{Verification and Results}
% What data will there be to analyze?
% How will that data be analyzed?
% Why is this data useful? What do we hope to get from it?

\section*{Other Risks and Discussion topics}
% What would we do if we can't get a teacher on board with us?
% How might we need to adjust the study if we cannot meet face to face because of COVID-19

% Research Objectives
\subsection*{Hypothesis}
\subsection*{Research Objectives}
\subsection*{Research Questions}

Determine if visual programming can be used to teach third grade students concepts in mathematics.
Does use of visual programming assist students in transitional learning?
In what ways does programming, and observing the execution of said program help students understand mathematical concepts?
Target audience
Students, educators, computer science community
Design of language
A scaffolded visual programming language.

Guided learning
Transitions through “levels” with Increasing complexity
Follows pedagogical practices of learning objectives,  with activities that can be adjusted to support those learning goals.

Programming blocks can be limited (you only get 2 or 10 or you can have as many as you need) This could be useful in providing hints to students.

Adaptive programming blocks that can be customized (show this field, don’t show this field) to the learning objectives.

\subsection*{Curriculum design}
\subsection*{Classroom design}
Which one would help us more effectively?
Individual option
Group option
Entire class option
How to get the data
	Pre-test
	Post-test
	Thinking out loud / ethnography
		Record the students using the program: observe their interaction with the program (finger movements, facial expressions, verbal interactions, interactions with classmates)
% How to analyze data
\subsection*{Data Analysis}
Improvement in pre and post-test
T-test
We don’t expect statistical significance (Why?) the study is too small 
Qualitative assessment (what methods we will use, interview, thinking out loud, etc)
\section*{Conclusion}
% What is a big impact? 
\subsection*{Implications}
Students will learn math
Computational thinking in real-world situations
What would happen if the study doesn’t work

Shortcomings that might lead to a failed state
What happens if we don’t find a teacher
Other risks needed to mitigate
COVID 19 updates (how would this look online?)

\bibliographystyle{plain}
\bibliography{thesis}

\end{document}
